Type-safe client-server communication in functional web applications

### 2014-03-01 paper.tex
appMain -> clientMain; remove explanation for Haste.App.

Our proposed model, dubbed ``Haste.App'' in an ingenious attempt at
capitalizing on the recent popularity of the ``app'' buzzword, has the
following properties:

@@

\begin{code}
appMain :: App Done
appMain = do
  greetings <- export helloServer

@@

\begin{code}
main :: IO ()
main =
  runApp (mkConfig "ws://localhost:1111" 1111) appMain
\end{code}

@@

Like \lstinline!runClient!, \lstinline!runApp! does not return, and is intended
to be the only computation executed by \lstinline!main!. In fact, performing
non-deterministic IO and allowing \lstinline!runApp! or its arguments to
capture the result may lead to unpredictable behavior. The \lstinline!appMain!
function should for all intents and purposes be regarded as the program's entry
point. A non-intrusive method of enforcing this using the type system is
described in section \ref{sec:limitations}, but is not used in the examples
given here as it relies on the use of the GHC plugin system.

@@

appMain :: App Done
appMain = do
  greetings <- export helloServer

@@

main :: IO ()
main =
  runApp (mkConfig "ws://localhost:1111" 1111) appMain
\end{code}
\caption{A seamless programming model: Hello Server}
\label{lst:hello-server}
\end{listingfloat}

@@

To allow the use of server state while treating the final argument to
\lstinline!runApp! (from here on always referred to as \lstinline!appMain!)
as the program's entry point, we may want to allow arbitrary IO computations to
be lifted into the \lstinline!App! monad:

@@

\begin{code}
appMain :: App Done
appMain = do
  recipients <- liftServerIO $ CC.newMVar []

@@

This is easily remedied however, by a slight compiler modification. Instead of
using \lstinline!main :: IO ()! as the program entry point, we introduce a
function \lstinline!appMain :: App Done! for this purpose, and let the runtime
system start the application by invoking \lstinline!runApp config appMain!,
where \lstinline!config! contains the URL and port configuration for the
server, given by the programmer using a compiler flag.

>>>>

Our proposed model, dubbed ``Haste.App'', has the following properties:

@@

\begin{code}
clientMain :: App Done
clientMain = do
  greetings <- export helloServer

@@

\begin{code}
main :: IO ()
main = runApp (mkConfig "ws://localhost:1111" 1111)
              clientMain
\end{code}

@@

Like \lstinline!runClient!, \lstinline!runApp! does not return, and is intended
to be the only computation executed by \lstinline!main!. In fact, performing
non-deterministic IO and allowing \lstinline!runApp! or its arguments to
capture the result may lead to unpredictable behavior. The \lstinline!clientMain!
function should for all intents and purposes be regarded as the program's entry
point. A non-intrusive method of enforcing this using the type system is
described in section \ref{sec:limitations}, but is not used in the examples
given here as it relies on the use of the GHC plugin system.

@@

clientMain :: App Done
clientMain = do
  greetings <- export helloServer

@@

main :: IO ()
main = runApp (mkConfig "ws://localhost:1111" 1111)
              clientMain
\end{code}
\caption{A seamless programming model: Hello Server}
\label{lst:hello-server}
\end{listingfloat}

@@

To allow the use of server state while treating the final argument to
\lstinline!runApp! (from here on always referred to as \lstinline!clientMain!)
as the program's entry point, we may want to allow arbitrary IO computations to
be lifted into the \lstinline!App! monad:

@@

\begin{code}
clientMain :: App Done
clientMain = do
  recipients <- liftServerIO $ CC.newMVar []

@@

This is easily remedied however, by a slight compiler modification. Instead of
using \lstinline!main :: IO ()! as the program entry point, we introduce a
function \lstinline!clientMain :: App Done! for this purpose, and let the runtime
system start the application by invoking \lstinline!runApp config clientMain!,
where \lstinline!config! contains the URL and port configuration for the
server, given by the programmer using a compiler flag.
